Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2021-01-06 papers/thesis-kirstie/thesis.tex
edits

In more recent times, other model fluids have been developed.
Zwanzig introduced the square-well 
fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading repulsive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures. 
In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
In this model, 
%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
hard spheres are given larger radii 
at low temperatures, and smaller radii at high temperatures. 
In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid. 
Models continue to be developed to characterize fluids with particular 
properties, or to address particular problems like surface tension, 
interfaces, adsorption, wetting or confinment.
\textcolor{red}{A review of model fluids, and problems studied more 
recently is given in ...}

While models for particular fluids were being created, powerful methods 
to predict the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
obtained~\cite{Hansen}. 
In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
of the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
%, and the corresponding equilibrium number density profile 
of a classical inhomogeneous fluid. 
Density functional theory was further developed for inhomogeneous classical fluids
by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium}.
To use classical Density Functional Theory (cDFT), an expression for the free energy 
as a functional of a spatially varying average single-particle number density 
%that varies spatially 
is constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency between results from SPT and Percus-Yevick, 
\textcolor{red}{In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
to form a functional for the hard-sphere fluid based on the idea that 
%the lowest term in the virial expansion of the excess chemical potential 
the lowest order term in the virial expansion of the excess chemical 
potential of a hard-sphere fluid can 
be expressed in terms of a weighted density using a weight function 
%using a weight 
%function 
of the form of 
a backward Heaviside step function}
%\begin{equation}
   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
%\end{equation}
that can itself be expressed in terms of a sum of convolutions involving 
a set of weight functions related to the geometry of a sphere. 
For a hard-sphere fluid with spheres of two different sized radii, 
$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
\end{equation}
FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
and for the direct correlation function approximation given by Percus-Yevick. 
However, it did not predict freezing due to one term in the expression that
diverged when the fluid was highly confined.  
Tarazona later solved this problem by introducing a tensor version of 
one of the energy density terms 
in the functional~\cite{tarazonaphi3, santos2012phi3} 
which took into account the non-spherical, and thus 
non-symmetric, shape of the cavity in which the fluid is confined 
while also yielding the correct excess free energy for the fluid as its 
confinment is reduced from a three-dimensional volume (as in a bulk fluid), 
to a two-dimensional plane (as in a thin-film), to a one-dimensional 
line (as in a fluid confined to a capillary tube), to a zero-dimensional 
dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.

In 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).
Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft, short-range repulsive and attractive potentials, and developed 
an error function model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
\textcolor{red}{For more recent developments in classical 
Density Functional Theory see (2020), Evan's review~\cite{evansreview} (2016), 
and Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013).}

FREEZING in literature :
\textcolor{red}{Classical density functional theory was used
to ... (paper from 2010), but not until recently has there been good success 
(other 23 papers...), and in all these the hard-sphere fluid has been 
used as the reference system.} 
In this paper, classical Density Functional 
Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to fit the 
functional to the soft-sphere WCA fluid.
%form the functional for the excess Helmholtz free energy. 
%for generating the weight functions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

Freezing of simple fluids is governed by entropy, which the universe
seeks to maximize according to the second law of Thermodynamics. When the 
entropy of a pure substance in its liquid state is lower than the entropy 
that it would have in its solid state at the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium), and the same chemical potential (necessary for diffusive 
equilibrium), a phase transition occurs, that is, the liquid 
freezes. 
Both the entropy and the volume (and thus the density) change abruptly 
at the phase transition.
The particular structure and density that result are affected 
by the interatomic potential. When repulsive, soft sphere fluids 
(that are not too soft) freeze, they form lattice structures composed of 
Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
This structure is what one would see by filling a box with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

@@

For pure substances, a phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure~\ref{fig:P-T_Diagram}. This phase diagram is 
representative of some arbitrary, unknown substance composed of atoms, 
or molecules which exhibit both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will just simply
be a fluid, which we will refer to as a liquid in this paper.
%which may be referred to as a "liquid", since at the phase boundary the fluid is dense. 
Points along the line that separates the solid and the liquid phases 
indicate a phase transition where the solid and liquid phases coexist, 
and show that both the solid and the liquid have the same temperature 
and the same pressure at a phase transition.

@@

\textcolor{blue}{
Rosenfeld begins the development of FMT by considering the lowest order 
term in the virial expansion of the excess chemical potential.
Taking the partial derivative of the free energy in the low density limit given 
by Equation~\ref{eq:Fexcess} gives
\begin{align}
   \mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
            &= k_BT\int -f(r)~n~d\vec r
\end{align}
For a hard-sphere fluid (simplified in this account by considering only 
equal sized spheres), the Mayer function, $f(r)$ can be written in terms 
of a backward Heaviside step function
\begin{align} 
    f(r) = -\theta(|\vec r|-2R)
\end{align}
where and $\vec r$ is the distance between the centers of two spheres.
The low-density excess free energy given by Equation~\ref{eq:Fexcess} 
can then be written as a weighted density.
\begin{align}
     F_{excess}=k_BT\left(\int n^2(\vec r)\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
\end{align} 
where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
Rosenfeld recognized that the Heaviside function describes mathmetically 
the exclusion volume of a pair of spheres, one of the two variables 
(the other one being the "overlap volume") in which the total 
excess chemical potential can be expanded. He also noted that the 
Heaviside function can uniquely be deconvolved into a set of weight functions
related to the geometry of a sphere.
\begin{align}
    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
\end{align} 
where again this expression is simplified in this account by considering only 
equal sized spheres. 
The excess free energy in the low density limit then takes the form
\begin{align}
     F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
\end{align}
%which can now be expressed in terms of a set of weighted densities
\begin{align}
     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
\end{align}
}
%In order to apply classical Density Functional Theory to a system that has
%a fixed temperature, volume, and number of particles, a functional for 
%the Helmholtz free energy must first be formed. An expression for the ideal 
%portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
%but forming an expression for the excess Helmholtz free
%energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
%being modeled can be challenging. 
%Soft Fundamental Measure Theory (SFMT) is a method for forming an 
%expression for the excess Helmholtz free energy of a soft-sphere fluid
%that does not depend on explicitly taking into account the interaction 
%that each sphere with every other sphere in the fluid (as required by 
%Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
%will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
%SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
%for soft-spheres. It is an extension of Fundamental Measure Theory (FMT) 
%originally developed by 
%Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
%for the hard-sphere fluid, and which expanded 
%Scaled Particle Theory (SPT)~\cite{santos2012phi3} 
%to inhomogeneous, or non-uniform hard-sphere fluids.

Scaled Particle Theory
seeks a general expression for the excess chemical 
potential of a homogeneous hard-sphere fluid which can then be integrated 
to find the excess Helmholtz free energy according to the relation
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogeneous fluid corresponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. 
\begin{align}
	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
\end{align}
Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemical potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation, shown below, is known as Widom's Insertion 
Formula~\cite{Hansen}.
\begin{align} \label{mu-der-int}
    \mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
\end{align}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
where 
\begin{align}
    <e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
\end{align}
and
\begin{align}
    Z=\sum_s{e^{-\beta V_s}}
\end{align}

>>>>

In more recent times, other model fluids have been developed.
Zwanzig introduced the square-well 
fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
a hard-edge attractive potential to the repulsive potential of the 
hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
spheres with softly fading repulsive potentials which he modeled by making
the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into 
each other when they collide at higher temperatures. 
In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres 
with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
In this model, 
%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
hard spheres are given larger radii 
at low temperatures, and smaller radii at high temperatures. 
In 1971 Weeks, 
Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive 
portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
which is known as the Weeks-Chandler-Anderson model fluid. 
Models continue to be developed to characterize fluids with particular 
properties, or to address particular problems like surface tension, 
interfaces, adsorption, wetting or confinment.
\textcolor{red}{A review of some model fluids and problems that have been
studied more recently is given by L\"{o}wen~\cite{lowen}.}

While models for particular fluids were being created, powerful methods 
to predict the properties of fluids were also being developed. 
In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
introduced from which an equation of state, and an expression for the 
excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
obtained~\cite{Hansen}. 
In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
of the hard-sphere fluid based on an approximate closure to the 
Ornstein-Zernike integral equation~\cite{PercusYevick}.
%Percus and Yevick took another approach to the hard-sphere fluid.
In 1961 Morita and Hiroike introduced a density functional 
approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
%, and the corresponding equilibrium number density profile 
of a classical inhomogeneous fluid. 
Density functional theory was further developed for inhomogeneous classical fluids
by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium}.
To use classical Density Functional Theory (cDFT), an expression for the free energy 
as a functional of a spatially varying average single-particle number density 
%that varies spatially 
is constructed for a particular fluid.
%DFT Evans 1979
%Led by the consistency between results from SPT and Percus-Yevick, 
\textcolor{red}{In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
to form a functional for the hard-sphere fluid based on the idea that 
%the lowest term in the virial expansion of the excess chemical potential 
the lowest order term in the virial expansion of the excess chemical 
potential of a hard-sphere fluid can 
be expressed in terms of a weighted density using a weight function 
%using a weight 
%function 
of the form of 
a backward Heaviside step function}
%\begin{equation}
   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
%\end{equation}
that can itself be expressed in terms of a sum of convolutions involving 
a set of weight functions related to the geometry of a sphere. 
For two spheres with radii, 
$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
\begin{displaymath}
   \theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
\end{equation}
FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
and for the direct correlation function approximation given by Percus-Yevick. 
However, it did not predict freezing due to one term in the expression that
diverged when the fluid was highly confined.  
Tarazona later solved this problem by introducing a tensor version of 
one of the energy density terms 
in the functional~\cite{tarazonaphi3, santos2012phi3} 
which took into account the non-spherical, and thus 
non-symmetric, shape of the cavity in which the fluid is confined 
while also yielding the correct excess free energy for the fluid as its 
confinment is reduced from a three-dimensional volume (as in a bulk fluid), 
to a two-dimensional plane (as in a thin-film), to a one-dimensional 
line (as in a fluid confined to a capillary tube), to a zero-dimensional 
dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.

In 1999 Schmidt adopted FMT for spheres with soft 
interatomic potential energies
(which rise to infinity over a distance, rather than instantaneously 
like the potential energy between hard spheres)
forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
%by replacing $w^{(3)}$ with a weight function that would result in  
%soft interatomic potentials (which rise to infinity over a distance, 
%rather than instantaneously).
Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft, short-range repulsive and attractive potentials, and developed 
an error function model 
%based on a Gaussian weight function 
for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.
%, similar to that exhibited by a WCA fluid. 
\textcolor{red}{For some more recent developments, see the review of new developments in 
classical Density Functional Theory by Evans ~\cite{evansreview} (2016), 
Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
and a recent generalization of the Barker-Henderson model to inhomogeneous 
fluids by Tschopp, Vuijk, Sharma and Brader~\cite{BHextended} (2020).}

\textcolor{red}{Classical density functional theory has also been applied 
to phase transitions, and freezing, but not until recently has there
been good progress. Ramakrishnan and Yussouff studied freezing by viewing a 
crystal as a highly inhomogeneous liquid~\cite{} (1979). Oettel and fellow 
researchers studied hard-sphere crystals with ~\cite{} (2010). Mortazavifar and Oettel
studied the fluid and crystal phases of the Asakura-Oosawa model fluid (2016). 
Lutsko and Lam studied studied crystallization of the Leonard-Jones fluid~\cite{lutsko} (2018),
and Lutsko and Schoonen looked at cDFT applied to the solid state~\cite{2020}.
In all these studies the hard-sphere fluid has been used as the reference system,
which is implemented with FMT, and to which other terms are added to create 
the full functional for the fluid being studied.}
In this paper, classical Density Functional Theory will be used to predict 
the freezing of a Weeks-Chandler-Anderson model fluid directly by applying 
SFMT with Schmidt's proposed error function model, 
in which we have incorporated our temperature dependent parameters to 
fit the functional to the soft-sphere WCA fluid. 
%form the functional for the excess Helmholtz free energy. 
%for generating the weight functions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

@@

%Freezing of simple fluids is governed by entropy, which the universe
%seeks to maximize according to the second law of Thermodynamics. When the 
%entropy of a pure substance in its liquid state is lower than the entropy 
%that it would have in its solid state at the same temperature (necessary 
%for thermal equilibrium with the environment), the same pressure (necessary for mechanical 
%equilibrium with the environment), and the same chemical potential (necessary for diffusive 
%equilibrium with the environment), a phase transition occurs, that is, the liquid 
%freezes. 
%%Along the phase boundary between pure liquid and pure solid states, both 
%%liquid and solid states can coexist in equilibrium. 
%Freezing of simple fluids can be seen when the entropy that a pure substance 
%would have in a liquid state is lower than the entropy 
%that it would have in its solid state at the same temperature (necessary 
%for thermal equilibrium), the same pressure (necessary for mechanical 
%equilibrium), and the same chemical potential (necessary for diffusive 
%equilibrium).

A liquid will eventually freeze as heat is removed, or the liquid is compressed. 
The state of a simple fluid can be characterized by its entropy per particle~\cite{entropyfingerprint}.
The entropy of a pure substance in a crystalline state is higher than the entropy 
that it would have in its liquid state at a given temperature and pressure, 
since there is more freedom for a particle to move between atoms  
at lattice sites than in a dense liquid with irregular and blocked pathways~\cite{Hansen}.
Both the entropy and the volume (and thus the density) change abruptly 
at the phase transition.
The particular structure and density that result are affected 
by the interatomic potential. When repulsive, soft sphere fluids 
(that are not too soft) freeze, they form lattice structures composed of 
Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
This structure is what one would see by filling a box with marbles. It 
results in the highest density achievable for hard, and nearly-hard, spheres 
that are purely repulsive.

@@

\textcolor{red}{A phase transition occurs along a line on a 
Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
diagram shown in Figure~\ref{fig:P-T_Diagram}. 
%
Points along the line that separates the solid and the liquid phases 
indicate a phase transition where the solid and liquid phases coexist, 
and show that both the solid and the liquid have the same temperature (necessary 
for thermal equilibrium)
and the same pressure (necessary for mechanical 
equilibrium)at a phase transition.
%
, and the same chemical potential (necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some arbitrary, unknown substance composed of atoms, 
or molecules which exhibit both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will just simply
be a fluid, which we will refer to as a liquid in this paper.}
%which may be referred to as a "liquid", since at the phase boundary the fluid is dense.

@@

\textcolor{blue}{
%Rosenfeld begins the development of FMT by considering the lowest order 
%term in the virial expansion of the excess chemical potential.
%Taking the partial derivative of the free energy in the low density limit given 
%by Equation~\ref{eq:Fexcess} gives
%\begin{align}
   %\mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
            %&= k_BT\int -f(r)~n~d\vec r
%\end{align}
Rosenfeld, who was studying mixtures, begins the development of FMT by considering that 
for a hard-sphere fluid the Mayer function, $f(r)$, can be written in terms 
of a backward Heaviside step function which, for a hard-sphere fluid 
composed of a mixture of spheres of radius $R_i$ and $R_j$, is
written as
\begin{align} 
    f(r) = -\theta(|\vec r_i - \vec r_j|-(R_i + R_j))
\end{align}
where and $|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
%
%The low-density excess free energy given by Equation~\ref{eq:Fexcess}
%The lowest order term of the excess chemical potential  
%can then be written as a weighted density.
%\begin{align}
     %\mu_{excess}=k_BT\int n\theta(|\vec r|-2R)~d\vec r
      %%F_{excess}=k_BT\left(\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
%\end{align} 
%where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
%The excess Helmholtz free energy given in Equation~\ref{eq:Fexcess} 
%can then be written as
%\begin{align}
      %F_{excess} = k_BTV\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r  
%\end{align} 
Rosenfeld recognized that the integral of the Heaviside step function
is equal to the volume of a sphere with a radius equal to that of the radii 
of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$
this becomes
\begin{align}
       \int_{Volume} \theta(|\vec r_i- \vec r_j|-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \\
       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + (4\pi R_j^2)R_i + \frac{4\pi}{3}R_j^3  \nonumber \\
       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)
\end{align}  
where the volume, surface area, radius, and the number 1 can be written as
\begin{align}
        V &= \int_{Volume} \theta(r-R)d\vec r \\
        S &= \int_{Volume} \delta(r-R) d\vec r \\
        R &= \int_{Volume} \frac{\delta(r-R)}{4\pi r} d\vec r \\
                1 &= \int_{Volume} \frac{\delta(r-R)}{4\pi r^2} d\vec r   
\end{align}
Limiting this discussion to a homogeneous hard-sphere fluid composed of hard spheres that have the same
radius R, the lowest order term of the virial expansion of the excess Helmholtz free energy which is given in Equation~\ref{eq:Fexcess} 
can now be written as
\begin{align}
      F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
                 &= k_BTn^2\left(\int \frac{\theta(|\vec r|-2R)}{2}~d\vec r\right)V  \\
                 &= k_BTn^2(2V(1) + 2SR)V  \\
                 &= k_BT\left[\left(\int n\theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
                 &= k_BT(n_3n_0 + n_2n_1)V \\
                 &= k_BT(\Phi_1 + \Phi_2)V \\
                 &= k_BT\Phi(\{n_i\})V
\end{align} 
where $n_i$ are weighted densities of the form
\begin{align}
     n_i=\int nw_i(r) d\vec r
\end{align}
with weight functions defined as
\begin{align} 
      w_3 &= \theta(r-R)  \label{scalor-weight3} \\
      w_2 &= \delta(r-R)  \\
      w_1 &= \frac{w_2}{4\pi r}  \\
      w_0 &= \frac{w_2}{4\pi r^2} \label{scalor-weight0}
\end{align}
and dimensionless energy densities $\Phi_1$ and $\Phi_2$ are defined as
\begin{align}
      \Phi_1 = n_3n_0  \\
      \Phi_2 = n_1n_2
\end{align}
The excess Helmholtz free energy can also be expressed in terms of the
packing fraction $\eta=nV$ as
\begin{align}
     F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
                &= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
                &= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
                &= k_BT4\eta^2
\end{align} 
%
For the inhomogeneous case where the number density varies spatially,
$n\rightarrow n(\vec r)$, Equation~\ref{eq:Fexcess} becomes
\begin{align}
     F_{excess} = k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}
\end{align}
where $\vec r_a$ is a position vector of one sphere, $\vec r_b$ is a position vector to another sphere, and 
$|\vec r_a - \vec r_b|$ is the distance between the centers of the two spheres.
Rosenfeld recognized that the Heaviside function can be written as 
a convolution of the weight functions, 
\begin{align}
    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
\end{align} 
where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
and the two vector weights are given by
\begin{align}
      \vec {w}_2 &= w_2~\hat r \\
      \vec {w}_1 &= w_1~\hat r
\end{align}
which are needed for the relation to be satisfied (although other expressions 
have been developed without using vector weights). 
%a uniquely deconvolved 
%into a set of weight functions related to the geometry of a sphere, and that it
%describes mathmetically 
%the exclusion volume of a pair of spheres, one of the two variables 
%(the other one being the "overlap volume") in which the total 
%excess chemical potential can be expanded. 
%
He then noted the following relation
\begin{align}
     w_\alpha(\vec r_a - \vec r_b) \otimes w_\gamma(\vec r_a - \vec r_b) &= \int w_\alpha(\vec r_a - \vec x) w_\gamma(\vec r_b - \vec x) d\vec x \\
                                                                         &= \int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x
\end{align}
which simplifies the equation for the excess Helmholtz free energy
\begin{align}
     F_{excess} &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}  \\
                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~(w_0\otimes w_3+ w_1 \otimes w_2+\vec {w}_1\otimes \vec {w}_2)  \\
                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)\left(\int w_3(\vec x - \vec r_a)w_0(\vec x - \vec r_b) \vec dx \color{white}\right)  \nonumber \\
                &+ \color{white}\left(\color{black}\int w_1(\vec x - \vec r_a)w_2(\vec x - \vec r_b) d\vec x  + \int \vec w_1(\vec x - \vec r_a)\cdot\vec w_2(\vec x - \vec r_b) d\vec x\right)    \\
                &= k_BT\int \left[\int n(\vec r_a) w_3(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_0(\vec x - \vec r_b)~d\vec r_b \color{white}\right]  \nonumber \\
                &+ \int n(\vec r_a) w_2(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_1(\vec x - \vec r_b)~d\vec r_b  \nonumber \\
                &+ \color{white}\left[\color{black} \int n(\vec r_a) \vec {w}_2(\vec x - \vec r_a)~d\vec r_a \cdot \int n(\vec r_b) \vec {w}_1(\vec x - \vec r_b)~d\vec r_b \right]\color{black}d\vec x  \\
                &= k_BT\int n_3n_0 + n_2n_1 + \vec n_1\cdot\vec n_2~d\vec x  \\
                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec x  \\
                & =k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r               
\end{align}
where $\vec n_1\cdot\vec n_2$=0 for the homogeneous case, 
and the weighted densities $n_i$ are defined as
\begin{align}
    n_i(\vec r)=\int n(\vec r')w_i(\vec r-\vec r')~d\vec r'
\end{align} 
%The excess free energy in the low density limit then takes the form
%\begin{align}
%    F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
%\end{align}
%%which can now be expressed in terms of a set of weighted densities
%\begin{align}
%     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
%\end{align}
%For a homogeneous fluid in the low density limit, the excess Helmholtz free
%energy can then be expressed as
%\begin{align}
      %F_{excess}=k_BTV\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r
%\end{align} 
%
%Rosenfeld arrived at an expression for the excess Helmholtz free energy 
%based on a set of weighted densities
%\begin{align}
    %F_{excess}=k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r
%\end{align} 
%
%The resulting expression for the excess Helmholtz free energy which is
%applicable to the more general case of an inhomogeneous fluid, reduces 
%to the same result obtained from Scaled Particle Theory for a homogeneous 
%hard-sphere fluid. 
}\textcolor{red}{
Lifting the restriction to the low density limit,
The expression for the excess Helmholtz free energy derived from FMT
is applicable to the more general case of an inhomogeneous fluid, but reduces 
to the same result obtained from Scaled Particle Theory for a homogeneous 
hard-sphere fluid. 
}
